Simulations of Ballot Polling Risk-Limiting Audits

### 2023-02-27 providence/sections/intro.tex
.

A total of 11, 315,056 votes were cast in the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.0566 in the pairwise contest. A pairwise contest between two candidates treats invalid votes and those received by other candidates as irrelevant to determining which of the two won the pairwise contest. Its {\em margin} is the difference between their vote counts as a fraction of the sum of their votes. Ballot polling audits used to audit government elections (such as \BRAVO, \Minerva and \Providence) audit a multi-candidate contest by conducting multiple audits of the pairwise contests between the winner and every other candidate.

@@

The second simulation assumed a tied election and did not stop at the end of five rounds. 
%The design of the \Providence audit (see section \ref{sec:prov}) ensures that the risk of the first round is no more than 0.9 times the risk limit, and the risk expended in the first round is roughly 0.09. 
Table \ref{tab:toy} lists the cumulative round sizes, cumulative votes for both candidates and the inverse \Providence ratio after each draw. Each round size was chosen for a conditional stopping probability of 0.9, given that the audit did not stop so far. For example, this corresponds to an approximate probability of 0.09 that the audit stops in the second round (that it does not stop in the first round and stops in the second), and so on. 
\begin{table}[h]
\centering
%\scriptsize
\begin{tabular}{|crrrc|}
\hline
\hline
Round& Cumulative& Trump & Biden & Inverse  \\
 No. & Round Size &  & & Ratio \\
\hline
\hline
1 & 2,217 & 1,111 & 1,079 & 0.256\\
2 & 5,970 & 2,940 & 2,953 & 4.251\\
3 & 16,685 & 8,281 & 8,171 & 3,150 \\
4 & 35,096 & 17,320 & 17,264 & 380,220,376 \\
5 & 76,979 & 37,943 & 37,868 & $2.5e^{+22}$ \\
\hline
\end{tabular}
\caption{A simulation of a \Providence audit of a tied contest with an announced margin of 0.0566, and $\alpha=0.1$. Each round size is computed for a conditional stopping probability of $0.9$. The stopping condition is Inverse Ratio < $\alpha$}
\label{tab:toy}
\end{table}

>>>>

A total of 11,315,056 votes were cast in the 2020 US President contest in the state of Texas \cite{MIT-Data-Lab}; candidate Trump won with 5,890,347 votes, and the second highest vote count was that of candidate Biden, who received 5,259,126 votes, for a margin of 0.0566 in the pairwise contest. A pairwise contest between two candidates treats invalid votes and those received by other candidates as irrelevant to determining which of the two won the pairwise contest. Its {\em margin} is the difference between their vote counts as a fraction of the sum of their votes. Ballot polling audits used to audit government elections (such as \BRAVO, \Minerva and \Providence) audit a multi-candidate contest by conducting multiple audits of the pairwise contests between the winner and every other candidate.

@@

The second simulation assumed a tied election and did not stop at the end of five rounds. 
%The design of the \Providence audit (see section \ref{sec:prov}) ensures that the risk of the first round is no more than 0.9 times the risk limit, and the risk expended in the first round is roughly 0.09. 
Table \ref{tab:toy} lists the cumulative round sizes, cumulative votes for both candidates and the inverse \Providence ratio after each draw. Each round size was chosen for a conditional stopping probability of 0.9, given that the audit did not stop so far. For example, this corresponds to an approximate probability of 0.09 that the audit stops in the second round (that it does not stop in the first round and stops in the second), and so on. 
\begin{table}[h]
\centering
%\scriptsize
\begin{tabular}{|crrrc|}
\hline
\hline
Round& Cumulative& Trump & Biden & Inverse  \\
 No. & Round Size &  & & Ratio \\
\hline
\hline
1 & 2,217 & 1,111 & 1,079 & 0.256\\
2 & 5,970 & 2,940 & 2,953 & 4.251\\
3 & 16,685 & 8,281 & 8,171 & 3,150 \\
4 & 35,096 & 17,320 & 17,264 & 380,220,376 \\
5 & 76,979 & 37,943 & 37,868 & $2.5e^{+22}$ \\
\hline
\end{tabular}
\caption{A simulation of a \Providence audit of a tied contest with an announced margin of 0.0566, and $\alpha=0.1$. Each round size is computed for a conditional stopping probability of $0.9$. The stopping condition is Inverse Ratio $\le \alpha$.}
\label{tab:toy}
\end{table}
